---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
---


```{r, message=FALSE, warning=FALSE, include=FALSE}

# -----------------------------------------------------------------------------------------------
# PDF analysis of pages in ICES advice reports.Rmd
#
# 11/01/2019 first coding; 
# -----------------------------------------------------------------------------------------------

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 
knit_hooks$set(crop = hook_pdfcrop)

# Libraries
library(pdftools)
library(tidyverse)
library(stringr)

# captioner library
library(captioner) # deal with captions
fig_nums <- captioner::captioner(prefix = "Figure ", levels=1, type=c("n"))

# Load utils code
source("../../mptools/r/my_utils.r")

# file path
my.path <- "E:/ICES/advice"

y <- 2018

# counter for loop
i <- 0

# loop over years
for (y in 2018:2018) {
  
  fl <- list.files(path=file.path(my.path, y), 
                   pattern = "pdf$", 
                   full.names = TRUE )
  print(y)
  
  for (fn in fl) {
    
    if (file.exists(fn) ) {
      # print(pdf_info(fn)[2])
      t <- 
        data.frame(fn=fn, year=y, npages = pdf_info(fn)[2], stringsAsFactors = FALSE) %>% 
        mutate(
          species = str_extract(basename(fn), "^[a-z]{3}"),
          species = ifelse(is.na(species), "ZZZ", species)       
        )
      if (i == 0) {acom_pages <- t} else {acom_pages <- bind_rows(acom_pages, t)}
      i <- i + 1
      
    } # end of file exists
  } # end of file loop
  
  gc()

} # end of y loop


# print(wg_pages)

# summarize by pentad of 5 years
acom_pages_bypentad <-
  acom_pages %>% 
  group_by(year) %>% 
  summarize(pages = sum(pages)) %>% 
  mutate(
    pentad = 10 / 2 * floor((year*2)/10)
  ) %>% 
  group_by(pentad) %>% 
  mutate(avg = mean(pages),
         end = pentad + 4) 


# Figure numbers
fig_nums(name    = "pages_byspecies", level = 1, display = FALSE,
         caption = "Number of pages by species")

fig_nums(name    = "pages_byyear", level = 1, display = FALSE,
         caption = "Number of pages by year and averages by pentad")

```



**Number of pages in advice documents**

M.A. Pastoors

`r format(Sys.time(), '%d/%m/%Y')`

Number of pages in advice document. Species have been automatically derived from the first three letters of the file name (and may not have worked flawlessly). A selection of species is shown in  
`r fig_nums("pages_byspecies", display="cite")`. All pages by 'species' with annual averages by decade is shown in `r fig_nums("pages_byyear", display="cite")`


```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

acom_pages %>% 
  filter(species %in% c("ang", "cap", "cod", "eel", "had", "her", "hke", "hom", "mac", "meg", "nep",
                        "ple", "san", "sol", "whg", "whb")) %>% 
  ggplot(aes(x=year,y=pages)) +
  theme_publication() +
  theme(legend.position = "none") +
  geom_bar(aes(fill=species), stat="identity") +
  expand_limits(y=0) +
  scale_x_continuous(breaks=scales::pretty_breaks()) +
  facet_wrap(~species)

```

*`r fig_nums("pages_byspecies")`*


```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

acom_pages %>% 

  ggplot(aes(x=year,y=pages)) +
  theme_publication() +
  theme(legend.title = element_blank()) +
  geom_bar(aes(fill=species), stat="identity") +
  geom_segment(data=acom_pages_bypentad, aes(x=pentad, xend=end, y=avg,yend=avg), colour="black", size=2) +
  geom_text   (data=acom_pages_bypentad, aes(x=pentad, y=avg, label=as.integer(avg)), 
               vjust=-0.5, hjust=0) +
  expand_limits(y=0) +
  guides(fill = guide_legend(nrow = 8)) + 
  scale_x_continuous(breaks=scales::pretty_breaks())


```

*`r fig_nums("pages_byyear")`*

```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

acom_pages %>% 
  group_by(year) %>% 
  summarize(nfiles = n()) %>% 
  
  ggplot(aes(x=year,y=nfiles)) +
  theme_publication() +
  theme(legend.title = element_blank()) +
  geom_bar(stat="identity") +
  expand_limits(y=0) +
  scale_x_continuous(breaks=scales::pretty_breaks())


```