---
output: 
  word_document:
    reference_docx: ../PFA_report_template_v1.5.dotx
---
```{r, message=FALSE, warning=FALSE, include=FALSE}
rm(list=ls())
```


```{r, message=FALSE, warning=FALSE, include=FALSE}

# -----------------------------------------------------------------------------------------------
# ICES advice and assessment summaries.Rmd
#
# 06/07/2018 first coding
# -----------------------------------------------------------------------------------------------

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=6) 
knit_hooks$set(crop = hook_pdfcrop)

# Libraries
library(rmarkdown) # creating markdown documents
library(tidyverse) # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
library(reshape2)  # reshaping data; e.g. dcast
library(ggthemes)  # for themes
library(pander)    # for print tables
library(readxl)    # read excel files
library(cowplot)   # multiplots
library(directlabels)  # label graphs
library(RColorBrewer) # colours
library(lubridate)
library(docxtractr) # read word documents
library(icesSD)    # ICES Stock database
library(icesSAG)   # ICES Stock Assessment Graphs

# Load utils code
source("../../mptools/r/my_utils.r")

# set dropboxdirs
assessdir <- paste(get_dropbox(), "/ICES assessment database", sep="")
advicedir <- paste(get_dropbox(), "/ICES advice database", sep="")

# =====================================================================================
# Load Stockkey and Rename data
# =====================================================================================

load(file=paste(assessdir, "/rdata/iStockkey.RData", sep=""))
load(file=paste(assessdir, "/rdata/iRename.RData", sep=""))

# read and process the iAdvice data
iAdvice <- 
  read_excel(paste0(advicedir, "/ICES Scientific Advice database 20180704.xlsx",sep=""), 
             sheet     = "DATA",
             col_names = TRUE,
             col_types = "text",
             skip      = 0)  %>% 
  lowcase() %>% 
  mutate_at( c("year", "assessyear", "ncpueseries","nsurveyseries"), 
             funs(as.integer)) %>% 
  mutate_at( c("advisedlandingsmax","advisedcatchmax","tal","tac",
               "officiallandings","iceslandings","icesindustrialbycatch","icesdiscards","icescatch",
               "fsqymin1","ssbymin1","fadvmax","fmax","f01","fmed","f35spr","flim","fpa","fmsy",
               "blim","bpa","msybtrig"), funs(as.numeric)) %>% 
  mutate( blim     = 1000 * blim, 
          msybtrig = 1000 * msybtrig) %>% 

  mutate_at( c("stockices","speciestype", "assessmodel", 
               "multispeciesmanagement", "stockmanagement", "adviceonstock", 
               "aggregated"), funs(tolower)) %>% 
  rename(stockkeylabelold = stockices, 
         assessmentyear = assessyear) %>% 
  arrange(stockkeylabelold, tacarea, year) %>%  
  mutate(species    = substr(stockkeylabelold,1,3) ) %>% 
  left_join(iStockkey, by="stockkeylabelold") %>% 
  
  filter(adviceonstock == "y")




# =====================================================================================
# Load Excel and QCS data
# =====================================================================================

excelqcs <-
  readxl::read_excel(
    path= paste(assessdir, "/data/QCS and EXCEL Assessment Database combined.xlsx", sep=""), 
    sheet = "data", col_names = TRUE, col_types = "text", trim_ws   = FALSE ) %>%
  lowcase() %>% 
  mutate_at(c("assessmentyear","year", 
              "lowrecruitment", "recruitment", "highrecruitment", "recruitmentage",
              "lowssb", "ssb", "highssb", 
              "lowf","f","highf", "flandings","fdiscards",
              "catches","discards", "landings"),   funs(as.numeric)) %>% 
  mutate_at(c("stockkey"),    funs(as.integer)) %>% 
  mutate_at(c("unitofrecruitment", "recruitmentdescription"),    funs(tolower)) %>% 
  mutate(recruitment       = ifelse(unitofrecruitment == "thousands", recruitment/1000, recruitment),
         unitofrecruitment = ifelse(unitofrecruitment == "thousands", "millions", unitofrecruitment),
         recruitment       = ifelse(unitofrecruitment == "billions", recruitment*1000, recruitment),
         unitofrecruitment = ifelse(unitofrecruitment == "billions", "millions", unitofrecruitment),
         source            = "excel") %>% 
  filter(purpose == "Advice")

excelqcs_unique <-
  excelqcs %>% 
  group_by(stockkey, stockkeylabelold, stockkeylabelnew, assessmentyear) %>% 
  filter(row_number()==1) %>% 
  select(stockkey, stockkeylabelold, stockkeylabelnew, assessmentyear) %>% 
  ungroup() 

# unique(excelqcs$purpose)
# unique(icesSAG$purpose)
# filter(excelqcs, stockkeylabelold == "bss-47") %>% View()

# =====================================================================================
# Load SAG database; also unpublished ones
# =====================================================================================

icesSAG <- 
  get(load(file=file.path(assessdir, "rdata/icesSAG 20180703.RData"))) %>% 
  lowcase() %>%
  rename(stockkeylabel = fishstock) %>%
  left_join(iRename, by="stockkeylabel") %>%
  left_join(iStockkey, by="stockkey") %>% 
  mutate_at(c("assessmentyear","year", 
              "lowrecruitment", "recruitment", "highrecruitment", "recruitmentage",
              "lowssb", "ssb", "highssb", 
              "lowf","f","highf", 
              "catches","discards", "landings"),   funs(as.numeric)) %>% 
  mutate(source = "sag") %>% 
  filter(purpose == "Advice")

icesSAG_unique <-
  icesSAG %>% 
  group_by(stockkey, stockkeylabelold, stockkeylabelnew, assessmentyear) %>% 
  filter(row_number()==1) %>% 
  select(stockkey, stockkeylabelold, stockkeylabelnew, assessmentyear) %>% 
  ungroup()


# =====================================================================================
# Merge SAG and excel database 
# =====================================================================================

iAssess <-
  
  # create excelqcs dataset to add
  dplyr::anti_join(select(excelqcs_unique, stockkey, stockkeylabelold, stockkeylabelnew, assessmentyear),
                   select(icesSAG_unique,  stockkey, stockkeylabelold, stockkeylabelnew, assessmentyear)) %>% 
  left_join(excelqcs, by=c("stockkey", "stockkeylabelold","stockkeylabelnew", "assessmentyear") ) %>% 
  data.frame() %>% 
  
  # add icesSAG
  bind_rows(., icesSAG)

# glimpse(excelqcs_unique)
# glimpse(icesSAG_unique)
# glimpse(iAssess)

```

```{r functions,  message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

# stock="her-47d3"
# firstassessyear=2014
# lastassessyear=2018
# firstyear=2000
# lastyear=2020

# Function to plot historical assessments
plot_history <- function(stock,firstassessyear,lastassessyear, firstyear, lastyear) {
  
  d <-
    iAssess %>% 
    ungroup() %>% 
    filter(stockkeylabelold == stock,
           assessmentyear   %in% c(firstassessyear:lastassessyear), 
           year             %in% firstyear:lastyear) %>% 
    
    mutate(tyear     = substr(as.character(assessmentyear), 3,4),
           type      = ifelse(assessmentyear == lastassessyear, "last", "other")) %>%
    select(assessmentyear, tyear, year, stockkey, stockkeylabel, stockkeylabelold, stockkeylabelnew,
           type, source,
           lowrecruitment, recruitment, highrecruitment, 
           lowf, f, highf, 
           lowssb, ssb, highssb) %>%
    data.frame()
  
  # plot ssb
  p1 <-
    d %>% 
    filter(!is.na(ssb)) %>%  

    ggplot(aes(year,ssb, group=tyear)) +
    
    theme_publication() +
    theme(legend.title=element_blank(),
          axis.text.x = element_text(angle = 0, vjust = 0.5, size=9),
          axis.text.y = element_text(size=9),
          # strip.background = element_blank(),
          legend.position = "null") +
    
    geom_ribbon(data=filter(d, type=="last"), 
                aes(x=year, ymin=lowssb, ymax=highssb, fill=type), 
                alpha=0.2) +
    geom_line(aes(colour = type, size=type, linetype=type) ) +
    
    geom_dl(aes(label  = tyear, colour = type), method = list(dl.combine("last.points"), cex = 0.8)) +
    
    geom_hline(data=filter(iAdvice, stockkeylabelold == stock, assessmentyear == lastassessyear),
               aes(x=assessmentyear, yintercept=blim, label="blim", vjust=-1), 
               colour="red", inherit.aes=FALSE) +
    geom_text (data=filter(iAdvice, stockkeylabelold == stock, assessmentyear == lastassessyear),
               aes( firstyear, blim, label = "blim", vjust = -0.3, hjust = 0),
               colour="red", inherit.aes = FALSE) +
    
    geom_hline(data=filter(iAdvice, stockkeylabelold == stock, assessmentyear == lastassessyear),
               aes(x=assessmentyear, yintercept=msybtrig), 
               colour="red", linetype="dashed", inherit.aes=FALSE) +
    geom_text (data=filter(iAdvice, stockkeylabelold == stock, assessmentyear == lastassessyear),
               aes( firstyear, msybtrig, label = "msy btrigger", vjust = -0.3, hjust = 0),
               colour="red", inherit.aes=FALSE) +
    
    scale_colour_manual  (values=c(last = "red",other="black")) +
    scale_fill_manual    (values=c(last = "red",other="black")) +
    scale_linetype_manual(values=c(last="solid",other="solid")) +
    scale_size_manual    (values=c(last= 1.0,   other=0.8)) +
    
    expand_limits(y = 0) +
    xlim(firstyear,lastyear) +
    labs(x = NULL, y = NULL , title = "SSB")  

  # F
  p2 <-
    d %>% 
    filter(!is.na(f)) %>%  

    ggplot(aes(year,f, group=tyear)) +
    
    theme_publication() +
    theme(legend.title=element_blank(),
          axis.text.x = element_text(angle = 0, vjust = 0.5, size=9),
          axis.text.y = element_text(size=9),
          # strip.background = element_blank(),
          legend.position = "null") +
    
    geom_ribbon(data=filter(d, type=="last"), 
                aes(x=year, ymin=lowf, ymax=highf, fill=type), 
                alpha=0.2) +
    geom_line(aes(colour = type, size=type, linetype=type) ) +
    
    geom_dl(aes(label  = tyear, colour = type), method = list(dl.combine("last.points"), cex = 0.8)) +

    geom_hline(data=filter(iAdvice, stockkeylabelold == stock, assessmentyear == lastassessyear),
               aes(x=assessmentyear, yintercept=fmsy), 
               colour="red", linetype="dashed", inherit.aes=FALSE) +
    geom_text (data=filter(iAdvice, stockkeylabelold == stock, assessmentyear == lastassessyear),
               aes( firstyear, fmsy, label = "fmsy", vjust = -0.3, hjust = 0),
               colour="red", inherit.aes=FALSE) +
    
    scale_colour_manual  (values=c(last = "red",other="black")) +
    scale_fill_manual    (values=c(last = "red",other="black")) +
    scale_linetype_manual(values=c(last="solid",other="solid")) +
    scale_size_manual    (values=c(last= 1.0,   other=0.8)) +
    
    expand_limits(y = 0) +
    xlim(firstyear,lastyear) +
    labs(x = NULL, y = NULL , title = "F")  

  # Recruitment
  p3 <-
    d %>% 
    filter(!is.na(recruitment)) %>%  

    ggplot(aes(year,recruitment, group=tyear)) +
    
    theme_publication() +
    theme(legend.title=element_blank(),
          axis.text.x = element_text(angle = 0, vjust = 0.5, size=9),
          axis.text.y = element_text(size=9),
          # strip.background = element_blank(),
          legend.position = "null") +
    
    geom_ribbon(data=filter(d, type=="last"), 
                aes(x=year, ymin=lowrecruitment, ymax=highrecruitment, fill=type), 
                alpha=0.2) +
    geom_line(aes(colour = type, size=type, linetype=type) ) +
    
    geom_dl(aes(label  = tyear, colour = type), method = list(dl.combine("last.points"), cex = 0.8)) +
    
    scale_colour_manual  (values=c(last = "red",other="black")) +
    scale_fill_manual    (values=c(last = "red",other="black")) +
    scale_linetype_manual(values=c(last="solid",other="solid")) +
    scale_size_manual    (values=c(last= 1.0,   other=0.8)) +
    
    expand_limits(y = 0) +
    xlim(firstyear,lastyear) +
    labs(x = NULL, y = NULL , title = "Recruitment")  

  # Combined plot
  # plot_grid(p1 + theme(legend.position  = "none", 
  #                    axis.title       = element_blank(),
  #                    strip.background = element_rect(colour=NA, fill=NA),
  #                    strip.text       = element_text(colour=NA)), 
  #         p2 + theme(axis.title       = element_blank()),
  #         p3 + theme(axis.title       = element_blank()),
  #         ncol=3, align = 'h')

    plot_grid(p1, p2,  ncol=2, align = 'hv') 
  
  } # End of plot_history function


# firstadviceyear <- 2016
# lastadviceyear  <- 2019
# stock           <- "her-47d3"

# Function to plot advice overviews
table_advice <- function(stock,firstadviceyear,lastadviceyear) {
  
  d <-
    iAdvice %>% 
    filter(stockkeylabelold == stock, 
           year %in% firstadviceyear:lastadviceyear) %>%
    mutate(advicebasis= substr(advicebasis,1,30),
           
           advice     = ifelse(!is.na(advisedcatch), advisedcatch, advisedlandings),
           advicemax  = ifelse(!is.na(advisedcatchmax), advisedcatchmax, advisedlandingsmax),
           advicetype = ifelse(!is.na(advisedcatch), "catch", "landings"),
           
           tac        = ifelse(!is.na(tac), tac, tal),
           tactype    = ifelse(!is.na(tac), "catch", "landings"), 
           
           catch      = ifelse(!is.na(icescatch), icescatch, NA),
           catchtype  = ifelse(!is.na(icescatch), "icescatch", ""),
           
           catchtype  = ifelse(!is.na(iceslandings) & is.na(catch), "iceslandings", catchtype),
           catch      = ifelse(!is.na(iceslandings) & is.na(catch), iceslandings, catch),
           
           catchtype  = ifelse(!is.na(officiallandings) & is.na(catch), "officiallandings", catchtype),
           catch      = ifelse(!is.na(officiallandings) & is.na(catch), officiallandings, catch), 
           
           advice_c   = 100 * (advicemax / lag(advicemax, n=1) - 1),
           tac_c      = 100 * (tac /lag(tac, n=1) - 1),
           catch_c    = 100 * (catch / lag(catch, n=1) -1)
           ) %>% 
    
    select(year, advicebasis, advice, fadvmax, advicemax, tac, catch, 
           advice_c, tac_c, catch_c,
           advicetype, tactype, catchtype) 
  
  d %>% 
  select(1, 2, 4, 5, 6, 7) %>%
  pandoc.table(., 
               style        = "simple",
               split.tables = 100, 
               justify      = "right",
               missing      =" ",
               big.mark     = '', 
               round        = c(0,0,2,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
}

table_advice2 <- function(stock,firstadviceyear,lastadviceyear) {
  
  iAdvice %>% 
    filter(stockkeylabelold == stock, 
           year %in% firstadviceyear:lastadviceyear) %>%
    mutate(advice     = ifelse(!is.na(advisedcatch), advisedcatch, advisedlandings),
           advicemax  = ifelse(!is.na(advisedcatchmax), advisedcatchmax, advisedlandingsmax),
           advicetype = ifelse(!is.na(advisedcatch), "catch", "landings"),
           
           tac        = ifelse(!is.na(tac), tac, tal),
           tactype    = ifelse(!is.na(tac), "catch", "landings"), 
           
           catch      = ifelse(!is.na(icescatch), icescatch, NA),
           catchtype  = ifelse(!is.na(icescatch), "icescatch", ""),
           
           catchtype  = ifelse(!is.na(iceslandings) & is.na(catch), "iceslandings", catchtype),
           catch      = ifelse(!is.na(iceslandings) & is.na(catch), iceslandings, catch),
           
           catchtype  = ifelse(!is.na(officiallandings) & is.na(catch), "officiallandings", catchtype),
           catch      = ifelse(!is.na(officiallandings) & is.na(catch), officiallandings, catch), 
           
           advice_c   = 100 * (advicemax / lag(advicemax, n=1) - 1),
           tac_c      = 100 * (tac /lag(tac, n=1) - 1),
           catch_c    = 100 * (catch / lag(catch, n=1) -1)
           ) %>% 
    
    select(year, advice_c, tac_c, catch_c, advicetype, tactype, catchtype) %>% 
    pandoc.table(., 
                 style        = "simple",
                 split.tables = 100, 
                 split.cells  = c(rep(7,10)),
                 justify      = "right",
                 missing      =" ",
                 big.mark     = '', 
                 round        = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
}

# table_advice(stock="her-47d3",firstadviceyear=2015,lastadviceyear=2019)

```


**Summary of ICES advice and assessments**

*June 2018 advice round*

Martin Pastoors, `r format(Sys.time(), '%d/%m/%Y')`

<!--Section 1: introduction ------------------------------------------------------------------- -->

# Introduction

In this document an overview is presented of the ICES assessment and advice for the most recent number of years for the stocks that are relevant to the PFA and Dutch fisheries in general.  

<!--Section 2: material and methods ----------------------------------------------------------- -->

# Material and methods

The overviews are based on two databases that have been developed as part of a more comprehensive research effort into the role of fisheries advice in European decision-making: one database containing the history of stock assessments and one database containing the history of scientific advice and management decisions.

<!--Section 3: advice and assessment overviews ------------------------------------------------ -->

# Assessment and advice overviews

## North Sea herring

```{r, echo=FALSE, message=FALSE, warning=FALSE}

s   <- "her-47d3"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)

```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}

plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}

table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

Assessment benchmarked in 2018. New natural mortality estimates lead to higher SSB and lower F. New method to standardize natural mortality estimates (no more scaling revision in future?). Revision of reference points. Weak 2014 year class leads to substantially lower advice, which was already foreseen in 2017. 
[her.27.3a47d.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.3a47d.pdf)



## Western Baltic Spring Spawning herring

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "her-3a22"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

Assessment benchmarked in 2018. Revision of reference points. Redefinition of Blim from 90 kT to 120 kT which leads to stock having been below Blim for a number of years already. No possibility of rebuilding stock to Blim within one year (which is the unwritten new ICES rule), and therefore a zero catch advice. An additional analysis will be undertaken (by MP) to evaluate alternative options for the zero catch advice. 
[her.27.20-24.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.20-24.pdf)



## Baltic Sea herring

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "her-2532-gor"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

Substantially lower estimate of SSB and higher estimate of fishing mortality in the most recent assessment is probably due to the variability in the survey index. The 2014 year class is estimated to be strong.  
[her.27.25-2932.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.25-2932.pdf)


## Herring in 6a and 7bc

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "her-67bc"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

The advice for a zero TAC is the same as in the previous years. Unfortunately ICES has not presented advice on the proposed rebuilding plan that has been developed by the PELAC and that has been submitted for review to ICES and that had been evaluated by the herring working group in March 2018. In previous years the monitoring TAC for herring in 6a and 7bc had been suggested by ICES as a means to maintain the knowledge base for this herring stock. The monitoring TAC is still mentioned under issues relevant for the advice. 
[her.27.6a7bc.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.6a7bc.pdf)


## Herring in the Irish Sea

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "her-nirs"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

Stock was benchmarked in 2017. SSB is well above MSY Btrigger and fishing mortality well below Fmsy. Potential mixing of herring in the Irish Sea and the Celtic Sea. 
[her.27.nirs.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.nirs.pdf)



## Herring in the Celtic Sea

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "her-irls"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

Stock went through interbenchmark in 2018 however perception of the stock trends is very similar to last year. Evaluations conducted in 2018 showed that the agreed management strategy is no longer precautionary. Therefore the advice is according to the MSY approach. 
[her.27.irls.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.irls.pdf)



## Sprat in the North Sea

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "spr-nsea"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

The change in recruitment level between this year’s assessment and the previous years is caused by a recent update of natural mortality of sprat in the first six months of their life. The 2018 assessment also revised the 2017 SSB downwards. The assessment has a tendency to overestimate above-average recruitments.
[spr.27.4.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/spr.27.4.pdf)



## Sprat in the Baltic Sea

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "spr-2232"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

Advice according to the Baltic Multiannual Plan (MAP). 
[spr.27.22-32.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/spr.27.22-32.pdf)



## Greater argentine in 5b and 6a

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "arg-5b6a"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Last assessment was in 2017. Biennial assessment based on survey trends")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

Latest advice was from 2017, applicable for two years. Next advice foreseen in 2019. 
[aru.27.5b6a.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2017/2017/aru.27.5b6a.pdf)




## Northern Hake

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "hke-nrtn"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

Stock size is well above MSY Btrigger and fishing mortality is below Fmsy. In the past, there has been a tendency to underestimate SSB and overestimate F but this is not the case in the most recent years. 
[hke.27.3a46-8abd.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/hke.27.3a46-8abd.pdf)



## North Sea cod

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "cod-347d"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

The 2018 assessment gave substantially lower estimates of SSB and higher estimates of fishing mortality in recent years. According to the advice document, this is (partially) caused by a lower than expected survey index for the strong 2013 and 2016 year classes in IBTS quarter 1 survey and the IBTS quarter 3 survey. The 2017 SSB is 32% lower compared to the estimate that was presented last year. This also leads to a substantially lower advice for 2019 compared to the advice for 2018. 
[cod.27.47d20.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/cod.27.47d20.pdf)



## North Sea plaice

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "ple-nsea"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

Stock size is well above MSY Btrigger (and close to 1 million tonnes) and fishing mortality is around Fmsy. An EU multiannual management plan (MAP) has been proposed for this stock (EU, 2016). This plan is not adopted by Norway, thus, not used as the basis of the advice for this shared stock. The advice is based on the MSY approach. 
[ple.27.420.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/ple.27.420.pdf)



## North Sea sole

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "sol-nsea"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

Stock size above MSY Btrigger, fishing mortality around Fmsy. Advice according to the EU multiannual plan (MAP) for the North Sea. 
[sol.27.4.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/sol.27.4.pdf)



## Atlantic Seabass

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "bss-47"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

ICES has reissued the advice for 2018 and issued an advice for 2019. ICES was requested by the EU to provide updated advice on fishing opportunities for 2018 because a benchmark of the assessment had been conducted in early 2018. The 2018 benchmark has not changed the perception of the current SSB (B below Blim). However, given the current low fishing mortality and the above-average recruitment in 2013 and 2014, SSB in 2019 and 2020 is predicted to be above Blim, allowing for some catches in 2018 and 2019. 
[bss.27.4bc7ad-h.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/bss.27.4bc7ad-h.pdf)


# More information

Please refer to the [ICES website](http://www.ices.dk/community/advisory-process/Pages/Latest-advice.aspx) for an overview of all advice documents. 